---
title: "CITE-seq Data Analysis"
author: "Lin Tian"
date: "5/31/2024"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Loading Library

```{r}
setwd("~/CITE_seq")
library(Seurat)
library(ggplot2) 
library(reshape2)
library(dplyr)
library(grid)
library(tidyverse)
```


## Filtering

```{r}
input <- Read10X(data.dir = "MC38-LvMet-KC_raw_matrix/")
data <- CreateSeuratObject(counts = input, project = "MC38_LvMet", min.cells = 5)
data[["percent.mt"]] <- PercentageFeatureSet(data, pattern = "^mt-")
data <- subset(data, subset=nFeature_RNA>200 & nFeature_RNA<4000 & percent.mt < 10)
data <- NormalizeData(data, normalization.method = "LogNormalize", scale.factor = 10000)
data <- FindVariableFeatures(data, selection.method = "vst", nfeatures = 2000)
all.genes <- rownames(data)
data <- ScaleData(data, features = all.genes)
data <- RunPCA(data, features = VariableFeatures(object = data))
DimPlot(data, reduction = "pca")
data <- JackStraw(data, num.replicate = 100)
data <- ScoreJackStraw(data, dims = 1:20)
plot1 <- JackStrawPlot(data, dims = 1:20)
plot2<-ElbowPlot(data)
CombinePlots(plots = list(plot1, plot2))
data <- FindNeighbors(data, dims = 1:20)
data <- FindClusters(data, resolution = 1.2)
data <- RunUMAP(data, dims = 1:20)
data <- RunTSNE(data, dims = 1:20)
pdf('umap.cluster.pdf')
DimPlot(data, reduction = "umap",label = F)
dev.off()
pdf('tsne.cluster.pdf')
DimPlot(data, reduction = "tsne",label = F)
dev.off()

for (l in levels (data)) {
  cluster1.markers <- FindMarkers(data, ident.1 = l, min.pct = 0.01,logfc.threshold = 0.25)
  cluster1.markers <- data.frame (gene = rownames (cluster1.markers))
  write.table (cluster1.markers,paste0("cluster.",l,".diffgenes.xls"),row.names = F,col.names = T, sep='\t',quote = F)
}

## adt  	
adtinputdir="MC38-LvMet-KC_filtered_citeseq_matrix/"
RAW.adt <- Read10X (data.dir = adtinputdir)
adt.names = rownames (RAW.adt)[32591:32718]
print (adt.names)
idx <- which(rownames (RAW.adt) %in% adt.names)
RAW.adt <- RAW.adt[idx,]
barcode <- gsub ("MC38_LvMet_","",rownames (data@meta.data))
adt <- RAW.adt[,barcode]
adt_assay <- CreateAssayObject (counts = adt)
PRO <- RenameCells (data,new.names = barcode)
PRO[["PROT"]] <- adt_assay
DefaultAssay (PRO) <- "RNA"
PRO <- RenameCells (PRO,add.cell.id = "MC38_LvMet")
DefaultAssay (PRO) <- "PROT"
PRO<- NormalizeData(PRO, normalization.method = 'CLR', margin = 2)
saveRDS (PRO,file.pat(outdir,"integrated_cluster.rds"))
```

## Load Seurat Object

```{r}
CITE_seq <- readRDS("CITE_seq.rds")
lvMet <- subset(CITE_seq, group%in%c("LvMet"))
DimPlot(lvMet,reduction = "umap",label = TRUE)
DefaultAssay(lvMet) <- 'RNA'
```

## Find variable genes for PCA

```{r}
lvMet <- FindVariableFeatures(object = lvMet,selection.method = "vst", nfeatures = 2000)
all.gene <- rownames(lvMet)
lvMet <- ScaleData(lvMet,features = all.gene) ## scale data
```

## Dimensional Reduction

```{r}
lvMet <- RunPCA(object=lvMet,features = VariableFeatures(object = lvMet))
DimPlot(lvMet, reduction = "pca")
ElbowPlot(lvMet,n=30)
```

## Cluster Cells

```{r}
lvMet <- FindNeighbors(lvMet, reduction = "pca", dims = 1:30)
lvMet <- FindClusters(lvMet,resolution = 1.5, algorithm = 1)
```

# Run non-linear dimensional reduction (tSNE and uMAP)

```{r}
lvMet <- RunTSNE(object=lvMet,dims.use=1:30,do.fast=TRUE,check_duplicates = FALSE)
lvMet <- RunUMAP(lvMet, reduction = "pca", dims = 1:30)
DimPlot(lvMet, reduction = "umap",group.by = "seurat_clusters",label = TRUE)
```

# Find markers

```{r}
markers <- FindAllMarkers(object = lvMet, 
                          only.pos = TRUE,
                          logfc.threshold= 0.25,
                          min.pct = 0.25)

top10 <- markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_log2FC)

DoHeatmap(object = lvMet, features=top10$gene)

write.table(markers,file="lvMet_dims30_resolution1.5.tsv",sep="\t") #write the markers of each cluster
```

## Determine cell types by markers according to The Human Protein Atlas and literature

```{r}
new.cluster.ids <- c("0"="Cxcl9+Mph", 
                     "1"="Ccl5+Mph", 
                     "2"="Ccl5+Mph", 
                     "3"="Spp1+Mph", 
                     "4"="Ccl5+Mph", 
                     "5"="Ccl5+Mph", 
                     "6"="Spp1+Mph", 
                     "7"="Cxcl9+Mph", 
                     "8"="Monocyte", 
                     "9"="Ccl5+Mph", 
                     "10"="Spp1+Mph", 
                     "11"="Vcan+Mph", 
                     "12"="Monocyte", 
                     "13"="Dendritic Cells", 
                     "14"="Stmn1+Mph", 
                     "15"="Neutrophil", 
                     "16"="Stmn1+Mph", 
                     "17"="Spp1+Mph",
                     "18"="KCs", 
                     "19"="Cxcl9+Mph", 
                     "20"="Dendritic Cells",
                     "21"="Dendritic Cells"
)
```

# Add cell type to the metadata as a new column

```{r}
lvMet <- RenameIdents(lvMet, new.cluster.ids)
lvMet$newtype <- lvMet@active.ident
DimPlot(lvMet, reduction = "umap",label = TRUE)
# saveRDS(lvMet,"lvMet.rds") ## This is the Seurat for liver metastasis-associated macrophages.
```

# Annotion CITE-seq object with metastatic macrophage sub-clusters

```{r}
meta <- CITE_seq@meta.data
meta$cell_name <- rownames(meta)
meta$cluster <- as.character(CITE_seq@active.ident)
ident <- meta$cluster
names(ident) <- rownames(meta)

ident1 <- as.character(lvMet@active.ident)
names(ident1) <- names(lvMet@active.ident)
for(i in names(ident1)){
  if( i %in% meta$cell_name){
    ident[i] <- ident1[i]}
}
meta$cluster <- ident
CITE_seq@meta.data <- meta
Idents(CITE_seq) <-'cluster'
```

## Visualize Annotated Seurat Object

```{r}
CITE_seq <- subset(CITE_seq, cluster%in%c("Ccl5+Mph","KCs","Spp1+Mph","Vcan+Mph","Stmn1+Mph","Cxcl9+Mph"))
p1 <- DimPlot(CITE_seq,reduction="umap", label = TRUE, cols=c("#386cb0", "#a6761d", "#33a02c", "#fdbf6f", "#f0027f", "#737373"))
pdf("CITE_seq_umap.pdf",height = 10,width = 11)
p1
dev.off()

saveRDS(CITE_seq,"CITE_seq_after.rds")
```


## Plot the markers of each type

```{r}
CITE_seq <- readRDS("~/CITE_seq/CITE_seq_after.rds")
FeaturePlot(CITE_seq, features=c("Spp1", "Cxcl9", "Ccr2", "Cx3cr1", "Cd68", "Clec4f"), cols=c("#d9d9d9", "#fbb4ae", "#e31a1c"))
FeaturePlot(CITE_seq, features=c("Spi1", "C1qc", "Ccr2", "Cx3cr1", "Spp1", "Trem2"), cols=c("#d9d9d9", "#e31a1c"))
markerdf2 <- read_tsv("ref_marker2.tsv")
markerdf2$gene <- as.character(markerdf2$gene)
vln.df <- as.data.frame(CITE_seq[["RNA"]]@data[markerdf2$gene,])
vln.df$gene <- rownames(vln.df)
vln.df <- melt(vln.df,id="gene")
colnames(vln.df)[c(2,3)] <- c("CB","exp")
CITE_seq$CB <- colnames(CITE_seq)
anno <- CITE_seq@meta.data[,c("CB","cluster")]
vln.df <- inner_join(vln.df,anno,by="CB")
vln.df$gene <- factor(vln.df$gene,levels = markerdf2$gene) 

vln.df%>%ggplot(aes(cluster,exp))+geom_violin(aes(fill=gene),scale = "width")+
  facet_grid(vln.df$gene~.,scales = "free_y")+
  scale_fill_brewer(palette = "Set3",direction = 1)+
  scale_x_discrete("")+scale_y_continuous("")+
  theme_bw()+
  theme(
    axis.text.x.bottom = element_text(angle = 45,hjust = 1,vjust = 1),
    panel.grid.major = element_blank(),panel.grid.minor = element_blank(),
    legend.position = "none"
  )

ggsave("marker_vln.pdf", width = 11, height = 22, units = "cm")
```

# Save the marker of each cluster

```{r}
Idents(CITE_seq) <-"cluster"
markers <- FindAllMarkers(object = CITE_seq, only.pos = TRUE, logfc.threshold= 0.25, min.pct = 0.25)

write.table(markers,file="CITE_seq_dims30_resolution1.5.tsv",sep="\t")
```

## Check Expression Levels for Proteins

```{r}
DefaultAssay (CITE_seq) <- "ADT"
pros <- rownames(CITE_seq)
pros[grep("Tim", pros)]
pros[grep("CD49", pros)]
pros[grep("CD11b", pros)]

FeaturePlot(CITE_seq, features=c("A0567-Tim-4", "A0078-CD49d", "A0014-CD11b", "A0563-CX3CR1"), cols=c("#d9d9d9", "#fbb4ae", "#e31a1c"))

FeaturePlot(CITE_seq, features=c("A0014-CD11b"), cols=c("#d9d9d9", "#e31a1c"))
VlnPlot(CITE_seq, features=c("A0905-CD107a-LAMP-1-"), group.by="group", pt.size = 0)
```







